---
title: "Atlanta Hawks Analytics Assessment"
author: "Cole R. Wagner"
date: "4/7/2021"
output:
  html_document:
    theme: "journal"
---

```{r include=FALSE, warning=FALSE, message=FALSE}
load("C:/Users/wagne/Downloads/PbP_18_19.Rda")
playerlist <- read.csv("C:/Users/wagne/Downloads/playerlist.csv")
library(tidyverse)
library(stargazer)
library(purrr)
library(kableExtra)
library(gridExtra)
```

### Part 1: *Data Comprehension*

#### (A)

```{r include=FALSE, warning=FALSE, message=FALSE}
pbp <- pbp %>%
  mutate(made_shot = case_when(SHOT_MADE == "True" ~ 1,
                               SHOT_MADE == "" ~ 0,
                               SHOT_MADE == "False" ~ 0),
         missed_shot = case_when(SHOT_MADE == "False" ~ 1,
                                 SHOT_MADE == "True" ~ 0,
                                 SHOT_MADE == "" ~ 0))

EMT <- sort(unique(pbp$EVENTMSGTYPE))

made_associations <- c()

for(i in EMT){
  pbp_EMT <- pbp %>%
    filter(EVENTMSGTYPE == i)
  made_associations <- c(made_associations, mean(pbp_EMT$made_shot))
}

missed_associations <- c()

for(i in EMT){
  pbp_EMT <- pbp %>%
    filter(EVENTMSGTYPE == i)
  missed_associations <- c(missed_associations, mean(pbp_EMT$missed_shot))
}

EMT[which(made_associations != 0)]

EMT[which(missed_associations != 0)]
```

a. Made shots are associated with an *EVENTMSGTYPE* value of **1**.

b. Missed shots are associated with an *EVENTMSGTYPE* value of **2**.

#### (B)

```{r message=FALSE, warning=FALSE, echo=FALSE}
elapsed_time_function <- function(n, elapsed){
  elapsed_last_score <- c()
  elapsed <- c(0, elapsed)
  
  for(i in 1:n){
    elapsed_last_score[i] <- elapsed[i + 1] - elapsed[i]
  }
  return(elapsed_last_score)
}


OT_adjuster_function <- function(period, time){
  for(i in 1:length(period)){
    if(period[i] == 6){
      time[i] <- time[i] + 300
    } else if(period[i] == 7){
      time[i] <- time[i] + 600
    } else if(period[i] == 8){
      time[i] <- time[i] + 900
    } else if(period[i] == 9){
      time[i] <- time[i] + 1200
    } else{
      time[i] <- time[i]
    }
  }
  return(time)
}

teams <- unique(c(levels(pbp$HOME_TEAM), levels(pbp$AWAY_TEAM)))

for(t in teams){
  team_home <- pbp %>%
    filter(HOME_TEAM == t)
  
  team_home <- team_home %>%
    group_by(GAME_ID) %>%
    nest()
  
  team_home <- team_home %>%
    mutate(data = map(.x = data, .f = ~ mutate(.x,
                                               TIME = OT_adjuster_function(period = .x$PERIOD, time = .x$TIME))))
  
  team_home <- team_home %>%
    mutate(data = map(.x = data, .f = ~ mutate(.x,
                                               time_elapsed = elapsed_time_function(n = length(.x$TIME), elapsed = .x$TIME),
                                               team_1 = .x$HOME_PLAYER_ID_1,
                                               team_2 = .x$HOME_PLAYER_ID_2,
                                               team_3 = .x$HOME_PLAYER_ID_3,
                                               team_4 = .x$HOME_PLAYER_ID_4,
                                               team_5 = .x$HOME_PLAYER_ID_5))) %>%
    unnest()
  
  team_away <- pbp %>%
    filter(AWAY_TEAM == t)
  
  team_away <- team_away %>%
    group_by(GAME_ID) %>%
    nest()
  
  team_away <- team_away %>%
    mutate(data = map(.x = data, .f = ~ mutate(.x,
                                               TIME = OT_adjuster_function(period = .x$PERIOD, time = .x$TIME))))
  
  team_away <- team_away %>%
    mutate(data = map(.x = data, .f = ~ mutate(.x,
                                               time_elapsed = elapsed_time_function(n = length(.x$TIME), elapsed = .x$TIME),
                                               team_1 = .x$AWAY_PLAYER_ID_1,
                                               team_2 = .x$AWAY_PLAYER_ID_2,
                                               team_3 = .x$AWAY_PLAYER_ID_3,
                                               team_4 = .x$AWAY_PLAYER_ID_4,
                                               team_5 = .x$AWAY_PLAYER_ID_5))) %>%
    unnest()
  
  team_all <- rbind.data.frame(team_home, team_away)
  
  team_lineups <- team_all %>%
    group_by(team_1, team_2, team_3, team_4, team_5) %>%
    nest()
  
  team_lineups <- team_lineups %>%
    mutate(seconds_played = map_dbl(.x = data, .f = ~ sum(.x$time_elapsed))) %>%
    arrange(desc(seconds_played))
  
  assign(paste(t, "lineups", sep = "_"), team_lineups)
  rm(team_away, team_home, team_all, team_lineups)
}

p76ers_lineups <- `76ers_lineups`
rm(`76ers_lineups`)

teams[1] <- "p76ers"
teams <- sort(teams)

Bucks_top <- Bucks_lineups[1, ]
Bulls_top <- Bulls_lineups[1, ]
Cavaliers_top <- Cavaliers_lineups[1, ]
Celtics_top <- Celtics_lineups[1, ]
Clippers_top <- Clippers_lineups[1, ]
Grizzlies_top <- Grizzlies_lineups[1, ]
Hawks_top <- Hawks_lineups[1, ]
Heat_top <- Heat_lineups[1, ]
Hornets_top <- Hornets_lineups[1, ]
Jazz_top <- Jazz_lineups[1, ]
Kings_top <- Kings_lineups[1, ]
Knicks_top <- Knicks_lineups[1, ]
Lakers_top <- Lakers_lineups[1, ]
Magic_top <- Magic_lineups[1, ]
Mavericks_top <- Mavericks_lineups[1, ]
Nets_top <- Nets_lineups[1, ]
Nuggets_top <- Nuggets_lineups[1, ]
p76ers_top <- p76ers_lineups[1, ]
Pacers_top <- Pacers_lineups[1, ]
Pelicans_top <- Pelicans_lineups[1, ]
Pistons_top <- Pistons_lineups[1, ]
Raptors_top <- Raptors_lineups[1, ]
Rockets_top <- Rockets_lineups[1, ]
Spurs_top <- Spurs_lineups[1, ]
Suns_top <- Suns_lineups[1, ]
Thunder_top <- Thunder_lineups[1, ]
Timberwolves_top <- Timberwolves_lineups[1, ]
TrailBlazers_top <- `Trail Blazers_lineups`[1, ]
Warriors_top <- Warriors_lineups[1, ]
Wizards_top <- Wizards_lineups[1, ]

lineups <- rbind.data.frame(Bucks_top,
                            Bulls_top,
                            Cavaliers_top,
                            Celtics_top,
                            Clippers_top,
                            Grizzlies_top,
                            Hawks_top,
                            Heat_top,
                            Hornets_top,
                            Jazz_top,
                            Kings_top,
                            Knicks_top,
                            Lakers_top,
                            Magic_top, 
                            Mavericks_top,
                            Nets_top,
                            Nuggets_top,
                            p76ers_top,
                            Pacers_top,
                            Pelicans_top,
                            Pistons_top,
                            Raptors_top,
                            Rockets_top,
                            Spurs_top,
                            Suns_top,
                            Thunder_top,
                            Timberwolves_top,
                            TrailBlazers_top,
                            Warriors_top,
                            Wizards_top)

lineups <- lineups %>%
  select(-data)

teams[18] <- "76ers"

lineups <- cbind(teams, lineups)

lineups[["team_1"]] <- playerlist[match(lineups[['team_1']], playerlist[['PERSON_ID']]), 'DISPLAY_FIRST_LAST']
lineups[["team_2"]] <- playerlist[match(lineups[['team_2']], playerlist[['PERSON_ID']]), 'DISPLAY_FIRST_LAST']
lineups[["team_3"]] <- playerlist[match(lineups[['team_3']], playerlist[['PERSON_ID']]), 'DISPLAY_FIRST_LAST']
lineups[["team_4"]] <- playerlist[match(lineups[['team_4']], playerlist[['PERSON_ID']]), 'DISPLAY_FIRST_LAST']
lineups[["team_5"]] <- playerlist[match(lineups[['team_5']], playerlist[['PERSON_ID']]), 'DISPLAY_FIRST_LAST']

lineups <- lineups %>%
  arrange(desc(seconds_played))

rank <- c(1:30)

lineups <- cbind(rank, lineups)

lineups %>%
  kable(format = "html", col.names = c("Rank", "Team", "Player 1", "Player 2", "Player 3", "Player 4", "Player 5", "Sec. Played"), caption = "Table 1: Most Common Lineup for Each NBA Team (2018-19)") %>%
  kable_styling("striped", font_size = 11) %>%
  add_footnote("The 'Sec. Played' column is incorrect due to erroneous home/away team classifications in the 2018-19 NBA play-by-play data. To abide by the time restriction, I could not correct the inaccurate home/away team classifications (however, the code would work properly with accurate data).", notation = "number")
```

#### (C)

```{r message=FALSE, warning=FALSE, echo=FALSE}
pbp_4 <- pbp %>%
  filter(PERIOD == 4)

pbp_4_3ft <- pbp_4 %>%
  filter(FREE_THROW_TOTAL == 3) %>%
  group_by(PLAYER1_NAME) %>%
  summarize(n = n() / 3) %>%
  arrange(desc(n))

rank <- as.character(c(1:10))

top10_4_3ft <- cbind.data.frame(rank, head(pbp_4_3ft, 10))

colnames(top10_4_3ft) <- c("Rank:", "3-Shot-Fouls-Drawn Leaders:", "N:")

pbp_4_1ft <- pbp_4 %>%
  filter(FREE_THROW_TOTAL == 1) %>%
  group_by(PLAYER1_NAME) %>%
  summarize(n = n()) %>%
  arrange(desc(n))

top10_4_1ft <- cbind.data.frame(rank, head(pbp_4_1ft, 10))

colnames(top10_4_1ft) <- c("Rank:","And-1-Fouls-Drawn Leaders:", "N:")

kable(list(top10_4_3ft, top10_4_1ft), format = "html", caption = "Table 2: NBA League Leaders in Fouls Drawn in the 4th Quarter (by Foul Type)") %>%
  kable_styling("striped", font_size = 11)
```

### Part 2: *Analytical Acumen* (Prompt 1)

```{r include=FALSE}
pbp <- pbp %>%
  mutate(home_made_shot = case_when(made_shot == 1 & grepl("PT", HOMEDESCRIPTION, fixed = TRUE) == 1 ~ 1,
                                    made_shot == 1 & grepl("PT", HOMEDESCRIPTION, fixed = TRUE) == 0 ~ 0,
                                    made_shot == 0 ~ 0),
         home_missed_shot = case_when(missed_shot == 1 & grepl("MISS", HOMEDESCRIPTION, fixed = TRUE) == 1 ~ 1,
                                    missed_shot == 1 & grepl("MISS", HOMEDESCRIPTION, fixed = TRUE) == 0 ~ 0,
                                    missed_shot == 0 ~ 0),
         home_turnover = case_when(TEAM_TURNOVER_COUNT > 0 & grepl("Turnover", HOMEDESCRIPTION, fixed = TRUE) == 1 ~ 1,
                                   TEAM_TURNOVER_COUNT > 0 & grepl("Turnover", HOMEDESCRIPTION, fixed = TRUE) == 0 ~ 0,
                                   is.na(TEAM_TURNOVER_COUNT) == 1 ~ 0),
         home_made_free_throw = case_when(FREE_THROW_MADE == "True" & grepl("Free Throw", HOMEDESCRIPTION, fixed = TRUE) == 1 ~ 1,
                                          FREE_THROW_MADE == "True" & grepl("Free Throw", HOMEDESCRIPTION, fixed = TRUE) == 0 ~ 0,
                                          FREE_THROW_MADE == "False" ~ 0,
                                          FREE_THROW_MADE == "" ~ 0),
         home_missed_free_throw = case_when(FREE_THROW_MADE == "False" & grepl("Free Throw", HOMEDESCRIPTION, fixed = TRUE) == 1 ~ 1,
                                            FREE_THROW_MADE == "False" & grepl("Free Throw", HOMEDESCRIPTION, fixed = TRUE) == 0 ~ 0,
                                            FREE_THROW_MADE == "True" ~ 0,
                                            FREE_THROW_MADE == "" ~ 0),
         home_off_rebound = case_when(REBOUND_OFFENSIVE_COUNT > 0 & grepl("REBOUND", HOMEDESCRIPTION, fixed = TRUE) == 1 ~ 1,
                                      REBOUND_OFFENSIVE_COUNT > 0 & grepl("REBOUND", HOMEDESCRIPTION, fixed = TRUE) == 0 ~ 0,
                                      REBOUND_OFFENSIVE_COUNT == 0 ~ 0,
                                      is.na(REBOUND_OFFENSIVE_COUNT) == 1 ~ 0),
         home_def_rebound = case_when(REBOUND_DEFENSIVE_COUNT > 0 & grepl("REBOUND", HOMEDESCRIPTION, fixed = TRUE) == 1 ~ 1,
                                      REBOUND_DEFENSIVE_COUNT > 0 & grepl("REBOUND", HOMEDESCRIPTION, fixed = TRUE) == 0 ~ 0,
                                      REBOUND_DEFENSIVE_COUNT == 0 ~ 0,
                                      is.na(REBOUND_DEFENSIVE_COUNT) == 1 ~ 0),
         home_three_shot_made = case_when(made_shot == 1 & grepl("3PT", HOMEDESCRIPTION, fixed = TRUE) == 1 ~ 1,
                                          made_shot == 1 & grepl("3PT", HOMEDESCRIPTION, fixed = TRUE) == 0 ~ 0,
                                          made_shot == 0 ~ 0),
         away_made_shot = case_when(made_shot == 1 & grepl("PT", VISITORDESCRIPTION, fixed = TRUE) == 1 ~ 1,
                                    made_shot == 1 & grepl("PT", VISITORDESCRIPTION, fixed = TRUE) == 0 ~ 0,
                                    made_shot == 0 ~ 0),
         away_missed_shot = case_when(missed_shot == 1 & grepl("MISS", VISITORDESCRIPTION, fixed = TRUE) == 1 ~ 1,
                                    missed_shot == 1 & grepl("MISS", VISITORDESCRIPTION, fixed = TRUE) == 0 ~ 0,
                                    missed_shot == 0 ~ 0),
         away_turnover = case_when(TEAM_TURNOVER_COUNT > 0 & grepl("Turnover", VISITORDESCRIPTION, fixed = TRUE) == 1 ~ 1,
                                   TEAM_TURNOVER_COUNT > 0 & grepl("Turnover", VISITORDESCRIPTION, fixed = TRUE) == 0 ~ 0,
                                   is.na(TEAM_TURNOVER_COUNT) == 1 ~ 0),
         away_made_free_throw = case_when(FREE_THROW_MADE == "True" & grepl("Free Throw", VISITORDESCRIPTION, fixed = TRUE) == 1 ~ 1,
                                          FREE_THROW_MADE == "True" & grepl("Free Throw", VISITORDESCRIPTION, fixed = TRUE) == 0 ~ 0,
                                          FREE_THROW_MADE == "False" ~ 0,
                                          FREE_THROW_MADE == "" ~ 0),
         away_missed_free_throw = case_when(FREE_THROW_MADE == "False" & grepl("Free Throw", VISITORDESCRIPTION, fixed = TRUE) == 1 ~ 1,
                                            FREE_THROW_MADE == "False" & grepl("Free Throw", VISITORDESCRIPTION, fixed = TRUE) == 0 ~ 0,
                                            FREE_THROW_MADE == "True" ~ 0,
                                            FREE_THROW_MADE == "" ~ 0),
         away_off_rebound = case_when(REBOUND_OFFENSIVE_COUNT > 0 & grepl("REBOUND", VISITORDESCRIPTION, fixed = TRUE) == 1 ~ 1,
                                      REBOUND_OFFENSIVE_COUNT > 0 & grepl("REBOUND", VISITORDESCRIPTION, fixed = TRUE) == 0 ~ 0,
                                      REBOUND_OFFENSIVE_COUNT == 0 ~ 0,
                                      is.na(REBOUND_OFFENSIVE_COUNT) == 1 ~ 0),
         away_def_rebound = case_when(REBOUND_DEFENSIVE_COUNT > 0 & grepl("REBOUND", VISITORDESCRIPTION, fixed = TRUE) == 1 ~ 1,
                                      REBOUND_DEFENSIVE_COUNT > 0 & grepl("REBOUND", VISITORDESCRIPTION, fixed = TRUE) == 0 ~ 0,
                                      REBOUND_DEFENSIVE_COUNT == 0 ~ 0,
                                      is.na(REBOUND_DEFENSIVE_COUNT) == 1 ~ 0),
         away_three_shot_made = case_when(made_shot == 1 & grepl("3PT", VISITORDESCRIPTION, fixed = TRUE) == 1 ~ 1,
                                          made_shot == 1 & grepl("3PT", VISITORDESCRIPTION, fixed = TRUE) == 0 ~ 0,
                                          made_shot == 0 ~ 0))

Hawks <- pbp %>%
  filter(HOME_TEAM == "Hawks" | AWAY_TEAM == "Hawks") %>%
  group_by(GAME_ID) %>%
  nest()

switchers <- c(5, 6, 10, 13, 15, 16, 18, 19, 20, 23, 24, 27, 30, 31, 32, 33, 38, 39, 40, 44, 45, 46, 48, 49, 51, 53, 54, 55, 57, 58, 59, 64, 65, 66, 67, 68, 69, 70, 78, 81, 82)

for(s in switchers){
  switch <- Hawks[[2]][[s]][[34]]
  Hawks[[2]][[s]][[34]] <- Hawks[[2]][[s]][[35]]
  Hawks[[2]][[s]][[35]] <- switch
  rm(switch)
}

Hawks <- Hawks %>%
  unnest()

Hawks_home <- Hawks %>%
  filter(HOME_TEAM == "Hawks") %>%
  group_by(GAME_ID) %>%
  nest()

Hawks_away <- Hawks %>%
  filter(AWAY_TEAM == "Hawks") %>%
  group_by(GAME_ID) %>%
  nest()

Hawks_home <- Hawks_home %>%
  mutate(final_margin = map_dbl(.x = data, .f = ~ as.numeric(as.character(.x$SCOREMARGIN[length(.x$SCOREMARGIN)]))))

Hawks_away <- Hawks_away %>%
  mutate(final_margin = map_dbl(.x = data, .f = ~ as.numeric(as.character(.x$SCOREMARGIN[length(.x$SCOREMARGIN)]))))

Hawks_home_wins <- Hawks_home %>%
  filter(final_margin > 0)

Hawks_home_losses <- Hawks_home %>%
  filter(final_margin < 0)

Hawks_away_wins <- Hawks_away %>%
  filter(final_margin < 0)

Hawks_away_losses <- Hawks_away %>%
  filter(final_margin > 0)

Hawks_home_wins <- Hawks_home_wins %>%
  mutate(EFG = map_dbl(.x = data, .f = ~ (sum(.x$home_made_shot) + 0.5 * sum(.x$home_three_shot_made)) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot))),
         OppEFG = map_dbl(.x = data, .f = ~ (sum(.x$away_made_shot) + 0.5 * sum(.x$away_three_shot_made)) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot))),
         TOV = map_dbl(.x = data, .f = ~ sum(.x$home_turnover) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot) + 0.44 * (sum(.x$home_made_free_throw) + sum(.x$home_missed_free_throw)) + sum(.x$home_turnover))),
         OppTOV = map_dbl(.x = data, .f = ~ sum(.x$away_turnover) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot) + 0.44 * (sum(.x$away_made_free_throw) + sum(.x$away_missed_free_throw)) + sum(.x$away_turnover))),
         ORB = map_dbl(.x = data, .f = ~ sum(.x$home_off_rebound) / (sum(.x$home_off_rebound) + sum(.x$away_def_rebound))),
         DRB = map_dbl(.x = data, .f = ~ sum(.x$home_def_rebound) / (sum(.x$home_def_rebound) + sum(.x$away_off_rebound))),
         FT = map_dbl(.x = data, .f = ~ sum(.x$home_made_free_throw) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot))),
         OppFT = map_dbl(.x = data, .f = ~ sum(.x$away_made_free_throw) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot))),
         Rate_3PT = map_dbl(.x = data, .f = ~ sum(.x$home_three_shot_made) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot))),
         OppRate_3PT = map_dbl(.x = data, .f = ~ sum(.x$away_three_shot_made) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot))),
         Rate_2PT = map_dbl(.x = data, .f = ~ (sum(.x$home_made_shot) - sum(.x$home_three_shot_made)) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot))),
         OppRate_2PT = map_dbl(.x = data, .f = ~ (sum(.x$away_made_shot) - sum(.x$away_three_shot_made)) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot))))


Hawks_home_losses <- Hawks_home_losses %>%
  mutate(EFG = map_dbl(.x = data, .f = ~ (sum(.x$home_made_shot) + 0.5 * sum(.x$home_three_shot_made)) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot))),
         OppEFG = map_dbl(.x = data, .f = ~ (sum(.x$away_made_shot) + 0.5 * sum(.x$away_three_shot_made)) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot))),
         TOV = map_dbl(.x = data, .f = ~ sum(.x$home_turnover) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot) + 0.44 * (sum(.x$home_made_free_throw) + sum(.x$home_missed_free_throw)) + sum(.x$home_turnover))),
         OppTOV = map_dbl(.x = data, .f = ~ sum(.x$away_turnover) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot) + 0.44 * (sum(.x$away_made_free_throw) + sum(.x$away_missed_free_throw)) + sum(.x$away_turnover))),
         ORB = map_dbl(.x = data, .f = ~ sum(.x$home_off_rebound) / (sum(.x$home_off_rebound) + sum(.x$away_def_rebound))),
         DRB = map_dbl(.x = data, .f = ~ sum(.x$home_def_rebound) / (sum(.x$home_def_rebound) + sum(.x$away_off_rebound))),
         FT = map_dbl(.x = data, .f = ~ sum(.x$home_made_free_throw) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot))),
         OppFT = map_dbl(.x = data, .f = ~ sum(.x$away_made_free_throw) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot))),
         Rate_3PT = map_dbl(.x = data, .f = ~ sum(.x$home_three_shot_made) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot))),
         OppRate_3PT = map_dbl(.x = data, .f = ~ sum(.x$away_three_shot_made) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot))),
         Rate_2PT = map_dbl(.x = data, .f = ~ (sum(.x$home_made_shot) - sum(.x$home_three_shot_made)) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot))),
         OppRate_2PT = map_dbl(.x = data, .f = ~ (sum(.x$away_made_shot) - sum(.x$away_three_shot_made)) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot))))


Hawks_away_wins <- Hawks_away_wins %>%
  mutate(EFG = map_dbl(.x = data, .f = ~ (sum(.x$away_made_shot) + 0.5 * sum(.x$away_three_shot_made)) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot))),
         OppEFG = map_dbl(.x = data, .f = ~ (sum(.x$home_made_shot) + 0.5 * sum(.x$home_three_shot_made)) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot))),
         TOV = map_dbl(.x = data, .f = ~ sum(.x$away_turnover) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot) + 0.44 * (sum(.x$away_made_free_throw) + sum(.x$away_missed_free_throw)) + sum(.x$away_turnover))),
         OppTOV = map_dbl(.x = data, .f = ~ sum(.x$home_turnover) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot) + 0.44 * (sum(.x$home_made_free_throw) + sum(.x$home_missed_free_throw)) + sum(.x$home_turnover))),
         ORB = map_dbl(.x = data, .f = ~ sum(.x$away_off_rebound) / (sum(.x$away_off_rebound) + sum(.x$home_def_rebound))),
         DRB = map_dbl(.x = data, .f = ~ sum(.x$away_def_rebound) / (sum(.x$away_def_rebound) + sum(.x$home_off_rebound))),
         FT = map_dbl(.x = data, .f = ~ sum(.x$away_made_free_throw) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot))),
         OppFT = map_dbl(.x = data, .f = ~ sum(.x$home_made_free_throw) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot))),
         Rate_3PT = map_dbl(.x = data, .f = ~ sum(.x$away_three_shot_made) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot))),
         OppRate_3PT = map_dbl(.x = data, .f = ~ sum(.x$home_three_shot_made) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot))),
         Rate_2PT = map_dbl(.x = data, .f = ~ (sum(.x$away_made_shot) - sum(.x$away_three_shot_made)) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot))),
         OppRate_2PT = map_dbl(.x = data, .f = ~ (sum(.x$home_made_shot) - sum(.x$home_three_shot_made)) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot))))


Hawks_away_losses <- Hawks_away_losses %>%
  mutate(EFG = map_dbl(.x = data, .f = ~ (sum(.x$away_made_shot) + 0.5 * sum(.x$away_three_shot_made)) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot))),
         OppEFG = map_dbl(.x = data, .f = ~ (sum(.x$home_made_shot) + 0.5 * sum(.x$home_three_shot_made)) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot))),
         TOV = map_dbl(.x = data, .f = ~ sum(.x$away_turnover) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot) + 0.44 * (sum(.x$away_made_free_throw) + sum(.x$away_missed_free_throw)) + sum(.x$away_turnover))),
         OppTOV = map_dbl(.x = data, .f = ~ sum(.x$home_turnover) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot) + 0.44 * (sum(.x$home_made_free_throw) + sum(.x$home_missed_free_throw)) + sum(.x$home_turnover))),
         ORB = map_dbl(.x = data, .f = ~ sum(.x$away_off_rebound) / (sum(.x$away_off_rebound) + sum(.x$home_def_rebound))),
         DRB = map_dbl(.x = data, .f = ~ sum(.x$away_def_rebound) / (sum(.x$away_def_rebound) + sum(.x$home_off_rebound))),
         FT = map_dbl(.x = data, .f = ~ sum(.x$away_made_free_throw) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot))),
         OppFT = map_dbl(.x = data, .f = ~ sum(.x$home_made_free_throw) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot))),
         Rate_3PT = map_dbl(.x = data, .f = ~ sum(.x$away_three_shot_made) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot))),
         OppRate_3PT = map_dbl(.x = data, .f = ~ sum(.x$home_three_shot_made) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot))),
         Rate_2PT = map_dbl(.x = data, .f = ~ (sum(.x$away_made_shot) - sum(.x$away_three_shot_made)) / (sum(.x$away_made_shot) + sum(.x$away_missed_shot))),
         OppRate_2PT = map_dbl(.x = data, .f = ~ (sum(.x$home_made_shot) - sum(.x$home_three_shot_made)) / (sum(.x$home_made_shot) + sum(.x$home_missed_shot))))

Hawks_wins <- rbind(Hawks_home_wins, Hawks_away_wins) %>%
  select(-data) %>%
  mutate(final_margin = abs(final_margin))

Hawks_losses <- rbind(Hawks_home_losses, Hawks_away_losses)%>%
  select(-data) %>%
  mutate(final_margin = -abs(final_margin))

Hawks <- rbind(Hawks_wins, Hawks_losses)

Hawks <- Hawks %>%
  mutate(win = case_when(final_margin > 0 ~ 1,
                         final_margin < 0 ~ 0))
```

```{css echo=FALSE}
table{
  width: 100%;
  font-size: 11px;
}
```

```{r results='asis', echo=FALSE}
Hawks_win_stats <- as.data.frame(Hawks_wins) %>%
  select(-GAME_ID)

Hawks_loss_stats <- as.data.frame(Hawks_losses) %>%
  select(-GAME_ID)

stargazer(Hawks_win_stats,
          type = "html",
          title = "Table 3: Atlanta Hawks Four Factors (by Wins vs. Losses)",
          summary.stat = c("n", "mean", "sd", "max", "min", "median"),
          covariate.labels = c("Final Score Margin (W)", "eFG% (W)", "Opp. eFG% (W)", "TOV% (W)", "Opp. TOV% (W)", "ORB% (W)", "DRB% (W)", "FT Rate (W)", "Opp. FT Rate (W)", "3PT Rate (W)", "Opp. 3PT Rate (W)", "2PT Rate (W)", "Opp. 2PT Rate (W)"),
          digits = 3)

stargazer(Hawks_loss_stats,
          type = "html",
          summary.stat = c("n", "mean", "sd", "max", "min", "median"),
          covariate.labels = c("Final Score Margin (L)", "eFG% (L)", "Opp. eFG% (L)", "TOV% (L)", "Opp. TOV% (L)", "ORB% (L)", "DRB% (L)", "FT Rate (L)", "Opp. FT Rate (L)", "3PT Rate (L)", "Opp. 3PT Rate (L)", "2PT Rate (L)", "Opp. 2PT Rate (L)"),
          digits = 3)
```

<br>

Intuitively, the Atlanta Hawks tend to have greater offensive production (as measured by eFG%, TOV%, ORB, FT Rate, etc.) and superior defensive production (as measured by Opp. eFG%, Opp. TOV%, DRB, Opp. FT Rate, etc.) in wins relative to losses. Two of these measures, however, warrant further discussion for their significant conditional mean disparities: effective field goal percentage (eFG%) and three-point rate (3PT Rate). According to Table 3, the Hawks' average eFG% in wins (55.7%) equates to 111% of their average eFG% in losses (50.4%). Similarly, the Hawks' average 3PT Rate in wins (15.9%) equates to 120% of their average 3PT Rate in losses (13.3%). Figures 1 and 2 provide further perspective below.

```{r warning=FALSE, message=FALSE, echo=FALSE, fig.width=10}
plot_1 <- ggplot(Hawks, aes(x = EFG, y = final_margin, color = OppEFG)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", color = "black") +
  scale_color_gradient(low = "gray80", high = "red") +
  theme_classic() +
  labs(x = "eFG%", y = "Final Score Margin", color = "Opp. eFG%", title = "Figure 1: Hawks eFG%")

plot_2 <- ggplot(Hawks, aes(x = Rate_3PT, y = Rate_2PT, color = as.factor(win))) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", color = "black") +
  scale_color_discrete(type = c("gray75", "red"), labels = c("Loss", "Win")) +
  theme_classic() +
  labs(x = "3PT Rate", y = "2PT Rate", color = "Game Result", title = "Figure 2: Hawks Shooting Trends")

grid.arrange(plot_1, plot_2, ncol = 2)
```

Figure 1 illustrates the positive relationship between eFG% and final score margin (Hawks' score - opponent's score). Cleverly, the plot also depicts the simultaneous dependence of the final score margin on Opp. eFG%, evident by the heavier concentration of darkly shaded points below the zero score margin and lightly shaded points above the zero score margin.

Figure 2 portrays the shooting-trend tradeoff between 2PT Rate (2P / FGA) and 3PT Rate (3P / FGA). The plot implies a negative relationship between shooting rates because teams must sacrifice 2P attempts to shoot more 3P attempts (hence, the "tradeoff" characterization). Furthermore, *holding 3PT Rate constant*, 2PT Rate exerts only a minor impact on win probability. On the other hand, *holding 2PT Rate constant*, 3PT Rate exerts a sizable impact on win probability, evidenced by the heavy concentration of red points (wins) on the right side of the plot (i.e., for higher values of 3PT Rate) at every fixed value of 2PT Rate.

Given the opportunity to expand this study, I would focus on the identification of optimal win predictors via logistic/probit regression. The summary statistics and graphics provide valuable information about team performance, but regression analysis could provide additional information related to the relative magnitudes of the effects of various (noncollinear) basketball metrics, which holds numerous applications in the evaluation of team quality. Additionally, I would investigate team performance variation in greater depth, which could provide valuable information related to performance consistency. For example, the examination of metric standard deviations (e.g., sd(eFG%)) could offer insight into the relationship between performance consistency and win probability.

```{r include=FALSE}
# References:

# https://stackoverflow.com/questions/27933076/how-can-i-add-notes-to-the-bottom-of-a-table-using-knitrkable

# https://stackoverflow.com/questions/1249548/side-by-side-plots-with-ggplot2

# https://stackoverflow.com/questions/37597126/searching-fields-using-grepl-in-r/37597302

# https://learn.datacamp.com/courses/introduction-to-data-visualization-with-ggplot2

# https://www.basketball-reference.com/teams/ATL/2019_games.html
```